Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-06-10 paper.Rnw
updates after updating limeaid (but something is not working with the explanation scatterplot)

# Join the no perms data (with poi removed) with the
# perms data (perms removed)
bullet_explain <- bullet_explain_noperms %>%
  anti_join(poi_cases_no_perms,
            by = c("implementation", "case")) %>%
  bind_rows(bullet_explain_perms %>%
              select(-perms_raw, -perms_numerified,
                     -perms_pred_simple, -perms_pred_complex,
                     -weights))

@@

# Create a metric comparison plot
plot_metrics(bullet_explain %>% mutate(label = as.factor(label)), add_lines = TRUE) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(color = "white", fill = "white"))
@
\caption{Plot of LIME assessment metrics for the 36 different applications of LIME applied to the bullet comparison data test set. The shape of the points represents the power used for the Gower distance calculation, and the color of the points represents the rank associated with that observation within a metric. The density simulation methods perform well for all metrics, but the bin based metrics often do not agree in terms of performance across metrics. \kgc{Something is wrong here. Right now, there are repetitions of some tuning parameter combinations.}}
\label{fig:bullet-assess-metric}
\end{figure*}

@@

eoi1_3qb <- plot_explain_scatter(bullet_explain_perms[1:3,], alpha = 0.9, weights = FALSE) +
  scale_color_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi2_3qb <- plot_explain_scatter(bullet_explain_perms[4:6,], alpha = 0.9, weights = FALSE) +
  scale_color_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi1_3eb <- plot_explain_scatter(bullet_explain_perms[7:9,], alpha = 0.9, weights = FALSE) +
  scale_color_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi2_3eb <- plot_explain_scatter(bullet_explain_perms[10:12,], alpha = 0.9, weights = FALSE) +
  scale_color_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi1_kd <- plot_explain_scatter(bullet_explain_perms[13:15,], alpha = 0.9, weights = FALSE) +
  scale_color_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi2_kd <- plot_explain_scatter(bullet_explain_perms[16:18,], alpha = 0.9, weights = FALSE) +
  scale_color_gradient2(low = "grey45",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  scale_fill_gradient2(low = "grey45",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

>>>>

# Join the no perms data (with poi removed) with the
# perms data (perms removed)
bullet_explain <- bullet_explain_noperms %>%
  anti_join(poi_cases_no_perms,
            by = c("implementation", "case")) %>%
  bind_rows(
    bullet_explain_perms %>%
      mutate(
        implementation = fct_recode(
          implementation,
          "2" = "1",
          "7" = "2",
          "31" = "3",
          "32" = "4"
        )
      ) %>%
      select(
        -perms_raw,
        -perms_numerified,-perms_pred_simple,
        -perms_pred_complex,-weights
      )
  )
@

@@

# Create a metric comparison plot
plot_metrics(bullet_explain %>% 
               mutate(label = as.factor(label)), 
             add_lines = TRUE) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(color = "white", fill = "white"))
@
\caption{Plot of LIME assessment metrics for the 36 different applications of LIME applied to the bullet comparison data test set. The shape of the points represents the power used for the Gower distance calculation, and the color of the points represents the rank associated with that observation within a metric. The density simulation methods perform well for all metrics, but the bin based metrics often do not agree in terms of performance across metrics.}
\label{fig:bullet-assess-metric}
\end{figure*}

@@

eoi1_3qb <- plot_explain_scatter(bullet_explain_perms[1:3,], alpha = 0.9, weights = FALSE) +
  scale_color_manual(values = c("grey50", "darkorange")) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi2_3qb <- plot_explain_scatter(bullet_explain_perms[4:6,], alpha = 0.9, weights = FALSE) +
  scale_color_manual(values = c("grey50", "darkorange")) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi1_3eb <- plot_explain_scatter(bullet_explain_perms[7:9,], alpha = 0.9, weights = FALSE) +
  scale_color_manual(values = c("grey50", "darkorange")) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi2_3eb <- plot_explain_scatter(bullet_explain_perms[10:12,], alpha = 0.9, weights = FALSE) +
  scale_color_manual(values = c("grey50", "darkorange")) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi1_kd <- plot_explain_scatter(bullet_explain_perms[13:15,], alpha = 0.9, weights = FALSE) +
  scale_color_manual(values = c("grey50", "darkorange")) +
  scale_fill_gradient2(low = "grey50",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))

eoi2_kd <- plot_explain_scatter(bullet_explain_perms[16:18,], alpha = 0.9, weights = FALSE) +
  scale_color_manual(values = c("grey50", "darkorange")) +
  scale_fill_gradient2(low = "grey45",
                        high = "darkorange",
                        midpoint = 0.5,
                        limits = c(0, 1)) +
  theme_bw(base_family = "Times", base_size = 8) +
  theme(strip.placement = "outside",
        strip.background = element_rect(color = "white",
                                        fill = "white"))
